---
title: "Milestone 3 — New pregame information and prospective edge discovery"
subtitle: "WR weekly PPR projections vs calibrated ESPN"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(ggplot2)
suppressMessages(tar_source("R"))
cfg <- tar_read(config)
fmt <- function(df, digits = 3) knitr::kable(mutate(df, across(where(is.double), ~ round(.x, digits))))
PAL <- c("#2a78d6", "#eb6834", "#1baf7a")
theme_report <- theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(), panel.grid.major = element_line(colour = "#e6e5e1", linewidth = 0.3),
        axis.text = element_text(colour = "#52514e"), plot.title.position = "plot",
        legend.position = "top", legend.justification = "left")
has_m3 <- file.exists("models/registry/m3.yml")
safe_read <- function(name) tryCatch(tar_read_raw(name), error = function(e) NULL)
```

## Summary

```{r summary-status, results = "asis"}
man <- readr::read_csv(PREDICTION_MANIFEST, col_types = readr::cols(.default = "c"))
ver <- verify_prediction_archive()
sv <- verify_snapshot_archive()
cat(sprintf("- Archived prospective runs: **%d** (weeks %s); hash-verified: **%s**.\n",
            nrow(man), paste(sort(unique(man$week)), collapse = ", "), all(ver$verified)))
cat(sprintf("- ESPN snapshots: **%d**; hash-verified: **%s**.\n", nrow(sv), all(sv$parquet_ok & sv$raw_ok)))
cat(sprintf("- Frozen lineages: %s.\n", paste(toupper(frozen_lineages()), collapse = ", ")))
```

## A. Did the M1/M2 prospective record remain intact?

```{r a-integrity}
man |>
  left_join(select(ver, run_id, verified), by = "run_id") |>
  transmute(run_id, week, lineages, n_rows, predicted_at_utc, first_kickoff_utc,
            snapshot_captured_at_utc, git_commit = substr(git_commit, 1, 7), verified) |>
  knitr::kable()
off <- official_predictions(ver)
st26 <- clean_player_stats(latest_live_file("player_stats", 2026), read_scoring_rules(cfg$scoring_system))
tg26 <- clean_team_games(latest_live_file("schedules", 2026))
pros <- score_prospective(off, st26, filter(tg26, week >= 5))
```

- **Completed prospective weeks: `r pros$weeks_complete`.**
- Frozen-model fingerprints, which must be identical:
  - M1: **`r tar_read(m1_fingerprint_check)$status`**
  - M2: **`r tar_read(m2_fingerprint_check)$status`**
  - M3: **`r if (has_m3) safe_read("m3_fingerprint_check")$status else "not frozen"`**

## B. What genuinely new information was investigated?

The full source feasibility matrix is in `docs/milestone3_plan.md` §13. In summary:

```{r b-matrix}
readr::read_csv("docs/source_feasibility.csv", show_col_types = FALSE) |> knitr::kable()
```

## C. Did new historical-valid information improve calibrated ESPN?

Only one new source passed the timing test with enough history to develop on:
**final weekly injury reports, 2017–2024**. Each row's own timestamp shows it
existed before kickoff. Two hypotheses were examined on development seasons
2020–2023 (experiment log E9):

- **C: ESPN under-reacts to vacated teammate targets.** **Rejected.** The
  residuals had the opposite sign and flipped across seasons, so no rule was frozen.
- **D: ESPN over-projects WRs designated Questionable.**
  - Development shortfall: −1.36 points (−18%), in every season.
  - Frozen as M3b (calibrated ESPN × 0.91 for Questionable WRs) **before** the
    one-time check on 2024, the only later season with timestamps.

```{r c-historical}
hb <- if (file.exists("models/registry/m3b.yml")) safe_read("m3b_holdout_comparisons") else NULL
if (!is.null(hb)) {
  hb |> select(model, subset, n, n_weeks, mae_diff, ci_low, ci_high, weeks_better, rmse_model, rmse_baseline) |> fmt(4)
}
```

In 2024 the Questionable rule improved calibrated ESPN with an interval excluding
0, better RMSE, and 89% of weeks won. It is small overall (about 0.3% of MAE),
because it touches only about 5% of WRs, but it is about 5% on those players.
This is **supporting, not confirmatory** evidence.

A second new source passed the timing test: the **git history of nflverse's
schedule file**. It is committed every 10–20 minutes from 2021, so it can be read
as of 2 hours before each kickoff.
- It provides pregame betting lines, which **resolves the closing-line problem**,
  and the **expected starting QB**, which matched the actual starter in 92.5% of
  team-games.
- **Hypothesis E: ESPN mis-projects WRs when the expected QB differs from the
  previous starter. Inconclusive, not frozen** (log E11). A backup starting showed
  −0.54 points (−9%), with a consistent sign in 2021–2023, but it is within noise,
  and the split was chosen after seeing the data.

## D. Which M3 prospective challengers were frozen?

```{r d-registry, results = "asis"}
for (lin in intersect(c("m3", "m3b"), frozen_lineages())) {
  reg <- read_registry(lin)
  cat(sprintf("\n**%s**, frozen at %s (parent commit `%s`):\n\n", toupper(lin), reg$frozen_at_utc,
              substr(reg$frozen_parent_commit, 1, 7)))
  purrr::iwalk(reg$models, function(m, id) {
    rule <- c(if (!is.null(m$role_rule)) sprintf("role: %s ≥ %s → %+.2f; ≤ %s → %+.2f", m$role_rule$metric,
                                                  m$role_rule$hi, m$role_rule$adj_hi, m$role_rule$lo, m$role_rule$adj_lo),
              if (!is.null(m$return_rule)) sprintf("return: 1 missed → %+.1f; 2+ missed → %+.1f",
                                                    m$return_rule$adj_one, m$return_rule$adj_two_plus),
              if (!is.null(m$questionable_rule)) sprintf("Questionable: %s %+.2f", m$questionable_rule$form,
                                                          m$questionable_rule$value))
    cat(sprintf("- **`%s`**: %s. %s\n", id, m$hypothesis, paste(rule, collapse = "; ")))
  })
  first_week <- man |> filter(grepl(paste0("(^|\\+)", lin, "($|\\+)"), lineages)) |> pull(week) |> as.integer()
  cat(sprintf("\nProspective record starts in **week %s**.\n",
              if (length(first_week)) min(first_week) else "—"))
}
```

One-time historical check of the frozen rules (2024–2025, rolling; not used to
change them):

```{r d-holdout, results = "asis"}
hc <- if (has_m3) safe_read("m3_holdout_comparisons") else NULL
if (!is.null(hc)) {
  print(knitr::kable(hc |> select(model, subset, n, n_weeks, mae_diff, ci_low, ci_high, weeks_better, rmse_model, rmse_baseline) |>
    mutate(across(where(is.double), ~ round(.x, 4)))))
  print(knitr::kable(safe_read("m3_holdout_seasonal") |> mutate(across(where(is.double), ~ round(.x, 4)))))
}
```

## E. What prospective evidence exists?

```{r e-prospective, results = "asis"}
if (pros$weeks_complete == 0) {
  cat("**Not enough prospective evidence yet.** No archived week has completed. Only archived, hash-verified",
      "predictions made before each game's kickoff will ever be scored. Nothing is back-filled.\n")
} else {
  print(knitr::kable(mutate(pros$metrics, across(where(is.double), ~ round(.x, 3)))))
  print(knitr::kable(mutate(pros$comparisons, across(where(is.double), ~ round(.x, 4)))))
}
```

## F. Did the role-change hypothesis work?

## G. Did the return-from-absence hypothesis work?

```{r fg, results = "asis"}
if (pros$weeks_complete < 8) {
  cat(sprintf("Prospectively: **insufficient evidence so far** (%d completed archived weeks; the pre-registered",
              pros$weeks_complete), "evaluation needs at least 8). The historical development evidence that",
      "motivated each rule, and the one-time 2024–2025 check above, are *not* confirmatory.\n")
}
```

## H. What did component modeling teach us?

```{r h-components}
cs <- tar_read(m3_component_summary)
cs$accuracy |> tidyr::pivot_wider(id_cols = component, names_from = level, values_from = mae) |> fmt()
cs$incremental |> select(component, mae_diff, ci_low, ci_high, weeks_better) |> fmt(4)
tar_read(m3_component_decomposition) |> fmt()
```

- Calibrated ESPN's points error comes from yards (about 23% of its variance),
  TDs (about 19%) and receptions (about 10%). The remainder is covariance between them.
- Our information adds a small, significant improvement only for **targets**,
  which measure opportunity. It **hurts** touchdown prediction.
- When we disagree with ESPN on a component, the realised error moves only about a
  third of the way toward us.

## I. Where does calibrated ESPN still fail?

```{r i-residuals, fig.height = 3.6, fig.width = 8}
ho <- tar_read(m2_holdout_aligned) |> filter(model == "m2_espn_cal")
ho |>
  mutate(bucket = cut(espn_proj, PROJECTION_BUCKETS, right = FALSE)) |>
  summarise(n = n(), mean_pred = mean(pred), mean_actual = mean(actual), mae = mean(abs(pred - actual)),
            large_miss = mean(abs(pred - actual) > 10), .by = bucket) |>
  arrange(bucket) |> fmt()
```

## J. Is there evidence of an incremental edge?

```{r j-verdict, results = "asis"}
cat(
  "**Prospectively: not enough evidence yet** (", pros$weeks_complete, " completed archived weeks).\n\n",
  "**Historically: promising but insufficient.**\n\n",
  "- A genuinely new pregame source, the timestamped Questionable designation, improved calibrated ESPN",
  "in the one later season that could check it (2024).\n",
  "- The pre-registered role and return rules held their direction in 2024–2025.\n",
  "- Every effect is small (about 0.1–0.3% of MAE overall).\n",
  "- Each rule was checked historically only once.\n\n",
  "Calibrated ESPN remains the bar. The 2026 prospective record of the frozen M3 and M3b rules is the",
  "test that can turn \"promising\" into \"yes\" or \"no\".\n", sep = " ")
```
