---
title: "Milestone 1 — Weekly WR projections"
subtitle: "Point-in-time dataset, benchmark, and chronological evaluation"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(ggplot2)

config <- tar_read(config)
audit <- tar_read(dataset_audit)
scoring_espn <- tar_read(scoring_validation)
leak <- tar_read(leakage_check)
ev_val <- tar_read(evaluation_validation)
selection <- tar_read(model_selection)
ev_test <- tar_read(evaluation_test)
headline <- tar_read(headline)
stats <- tar_read(player_stats)

has_espn <- identical(ev_val$population, "espn")
val_season <- config$splits$validation_season
test_season <- config$splits$test_season

# Categorical slots in fixed order (dataviz reference palette).
PAL <- c("#2a78d6", "#eb6834", "#1baf7a")
theme_report <- theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "#e6e5e1", linewidth = 0.3),
    axis.text = element_text(colour = "#52514e"),
    plot.title.position = "plot",
    legend.position = "top", legend.justification = "left"
  )
fmt <- function(df, digits = 3) {
  knitr::kable(mutate(df, across(where(is.double), ~ round(.x, digits))))
}
```

## Scope

- **Population:** ``r ev_val$population``.
  `espn` means WR player-weeks with an ESPN pregame projection above 0.
  `active` means WRs on the game-day active roster, the ESPN-free population.
- **Benchmark:** ``r ev_val$baseline``.
- **Splits:** train `r config$splits$first_train_season`–`r config$splits$train_last_season`,
  validation `r val_season`, test `r test_season` (untouched until the model was selected).
- **Scoring:** `r config$scoring_system`, computed from nflverse stats with our own tested code.

```{r espn-status, results = "asis"}
if (!has_espn) {
  cat("::: {.callout-warning}\n## ESPN benchmark not yet evaluated\n",
      "ESPN fetching is disabled pending a terms-of-use decision (see `docs/espn_projections.md`).",
      "Every result below compares our models to the **naive trailing-average baseline** on the",
      "active-roster population, **not to ESPN**. Nothing here shows that ESPN was beaten.\n:::\n")
}
```

## Data coverage

```{r coverage}
fmt(audit$by_season, 0)
```

Rows whose team had no game that week are excluded; there were
`r sum(audit$excluded_no_game$rows)` such rows in total.

## Scoring validation

```{r scoring}
off <- filter(stats, stats_position %in% c("QB", "RB", "WR", "TE"))
tibble(
  reference = "nflverse fantasy_points_ppr",
  player_weeks = nrow(off),
  share_identical = mean(abs(off$fantasy_pts - off$nflverse_ppr) < 0.011),
  n_different = sum(abs(off$fantasy_pts - off$nflverse_ppr) >= 0.011)
) |> fmt(4)
```

Differences from nflverse are intended. ESPN penalises fumbles lost on kick
and punt returns and credits offensive fumble-recovery TDs, while nflverse's
PPR column does neither. The authoritative check is against **ESPN's own actual
totals**:

```{r scoring-espn, results = "asis"}
if (isTRUE(scoring_espn$available)) {
  print(knitr::kable(mutate(scoring_espn$summary, across(where(is.double), ~ round(.x, 4)))))
  cat("\nRemaining mismatches (", nrow(scoring_espn$mismatches), "): special-teams fumble-recovery TDs",
      "by coverage players, a missed-field-goal return TD, and two 2-pt conversions that ESPN",
      "does not credit to the player.\n")
} else {
  cat("ESPN actual totals are unavailable (ESPN disabled), so validation against ESPN is pending.\n")
}
```

## Leakage checks

The empirical check corrupted every outcome at or after `r leak$n_cutoffs`
random cutoff weeks and recomputed that week's features.
**Leaking features found: `r leak$leaking_features`.**
The pipeline aborts if this is ever non-zero. The unit tests in
`tests/testthat/` cover windowing, as-of joins and season boundaries.

## Validation season (`r val_season`)

Candidates were fit on the training seasons. Under the rolling protocol they are
re-fit every week on all prior weeks. All models are compared on identical
player-weeks.

```{r validation}
ev_val$overall |>
  select(protocol, model, n, mae, rmse, bias, cor, rank_cor) |>
  arrange(protocol, mae) |>
  fmt()
```

**Selected by the pre-registered rule** (lowest rolling MAE, excluding
diagnostics): **``r selection$model``**.

## Test season (`r test_season`)

```{r test}
ev_test$overall |>
  select(protocol, model, n, mae, rmse, bias, cor, rank_cor) |>
  arrange(protocol, mae) |>
  fmt()
```

**Headline (rolling protocol):** paired, week-clustered bootstrap of
MAE(model) − MAE(``r headline$baseline``). Negative values favour the model.

```{r headline}
headline$bootstrap |>
  select(model, baseline, subset, n, n_weeks, mae_diff, ci_low, ci_high, weeks_model_better) |>
  fmt()
```

Pre-registered criterion: the CI must lie entirely below 0 **and** RMSE must not
be worse. Result: **`r if (headline$beats_baseline) "met" else "not met"`**
against ``r headline$baseline``.

### Is the result repeatable?

One season is not enough evidence. The table puts the same comparisons side by
side for the validation and test seasons (rolling protocol, all player-weeks).

```{r repeatability}
both <- bind_rows(
  mutate(ev_val$bootstrap, season = val_season),
  mutate(ev_test$bootstrap, season = test_season)
) |>
  filter(protocol == "rolling", subset == "all",
         model %in% c(selection$model, "espn_recal_l2", "espn_recal_l1", "ols_usage")) |>
  select(model, season, mae_diff, ci_low, ci_high, weeks_model_better) |>
  arrange(model, season)
fmt(both)
bias_tab <- bind_rows(ev_val$overall, ev_test$overall) |>
  filter(protocol == "rolling", model == headline$baseline) |>
  select(season, model, mae, rmse, bias)
fmt(bias_tab)
```

`espn_recal_l2` is a + b·ESPN fitted on past seasons. It contains **no new
information**: it only corrects ESPN's average level. When it captures most of
a model's gain, the gain is mainly calibration, such as ESPN over-projecting in
a given season, rather than better player-level information. `espn_recal_l1`
targets the median and shows how much MAE can be "won" through distributional
targeting alone. Its RMSE is worse.

## Error analysis

```{r weekly-mae, fig.height = 3.6, fig.width = 8}
show <- unique(c(if (has_espn) "espn", selection$model, "naive_roll8"))
wk <- ev_test$by_week |>
  filter(protocol == "rolling", model %in% show) |>
  mutate(model = factor(model, levels = show))
ggplot(wk, aes(week, mae, colour = model)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 2) +
  scale_colour_manual(values = PAL[seq_along(show)], name = NULL) +
  scale_x_continuous(breaks = seq(1, 18, 2)) +
  labs(title = paste("Weekly MAE,", test_season, "(rolling protocol)"),
       x = "Week", y = "MAE (PPR points)") +
  theme_report
```

```{r residuals, fig.height = 3.2, fig.width = 8}
res <- ev_test$residuals |> filter(protocol == "rolling") |> arrange(q50)
ggplot(res, aes(y = reorder(model, q50))) +
  geom_vline(xintercept = 0, colour = "#b5b4af", linewidth = 0.4) +
  geom_linerange(aes(xmin = q05, xmax = q95), colour = "#9ec5f4", linewidth = 0.7) +
  geom_linerange(aes(xmin = q25, xmax = q75), colour = PAL[1], linewidth = 2) +
  geom_point(aes(x = q50), colour = "#0d366b", size = 2.2) +
  labs(title = "Prediction error (prediction − actual), test season",
       subtitle = "Dot: median. Thick bar: 25th–75th percentile. Thin bar: 5th–95th.",
       x = "PPR points", y = NULL) +
  theme_report
```

The median error is positive while the mean error (bias) is close to 0.
Weekly WR scoring is right-skewed, so a forecast of expected points sits above
the typical (median) outcome. MAE is minimised by the conditional *median*, so
MAE alone can be improved simply by shrinking projections. This is why the
headline criterion also requires RMSE not to get worse.

```{r buckets, results = "asis"}
if (has_espn && nrow(ev_test$by_bucket) > 0) {
  cat("### By ESPN projection range\n\n")
  ev_test$by_bucket |>
    filter(protocol == "rolling", model %in% show) |>
    select(model, proj_bucket, n, mae, rmse, bias) |>
    arrange(proj_bucket, mae) |>
    fmt() |>
    print()
}
```

## Limitations

See `README.md` and `docs/` for the full list. In particular:

- `r if (!has_espn) "**No ESPN comparison has been run.**" else "ESPN projections were retrieved retroactively. Archived captures support that they are final pregame values (see docs/espn_projections.md)."`
- Closing betting lines are only valid for a kickoff-time prediction.
- Injury designations have no capture timestamp.
- One test season gives only 18 weekly clusters. Whether gains repeat across seasons still has to be shown with the 2026 season as a new holdout.
