---
title: "Milestone 3 — New pregame information and prospective edge discovery"
subtitle: "WR weekly PPR projections vs calibrated ESPN"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(ggplot2)
suppressMessages(tar_source("R"))
cfg <- tar_read(config)
fmt <- function(df, digits = 3) knitr::kable(mutate(df, across(where(is.double), ~ round(.x, digits))))
PAL <- c("#2a78d6", "#eb6834", "#1baf7a")
theme_report <- theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(), panel.grid.major = element_line(colour = "#e6e5e1", linewidth = 0.3),
        axis.text = element_text(colour = "#52514e"), plot.title.position = "plot",
        legend.position = "top", legend.justification = "left")
has_m3 <- file.exists("models/registry/m3.yml")
safe_read <- function(name) tryCatch(tar_read_raw(name), error = function(e) NULL)
```

## Summary

```{r summary-status, results = "asis"}
man <- readr::read_csv(PREDICTION_MANIFEST, col_types = readr::cols(.default = "c"))
ver <- verify_prediction_archive()
sv <- verify_snapshot_archive()
cat(sprintf("- Archived prospective runs: **%d** (weeks %s); hash-verified: **%s**.\n",
            nrow(man), paste(sort(unique(man$week)), collapse = ", "), all(ver$verified)))
cat(sprintf("- ESPN snapshots: **%d**; hash-verified: **%s**.\n", nrow(sv), all(sv$parquet_ok & sv$raw_ok)))
cat(sprintf("- Frozen lineages: %s.\n", paste(toupper(frozen_lineages()), collapse = ", ")))
```

## A. Did the M1/M2 prospective record remain intact?

```{r a-integrity}
man |>
  left_join(select(ver, run_id, verified), by = "run_id") |>
  transmute(run_id, week, lineages, n_rows, predicted_at_utc, first_kickoff_utc,
            snapshot_captured_at_utc, git_commit = substr(git_commit, 1, 7), verified) |>
  knitr::kable()
off <- official_predictions(ver)
st26 <- clean_player_stats(latest_live_file("player_stats", 2026), read_scoring_rules(cfg$scoring_system))
tg26 <- clean_team_games(latest_live_file("schedules", 2026))
pros <- score_prospective(off, st26, filter(tg26, week >= 5))
```

- **Completed prospective weeks: `r pros$weeks_complete`.**
- Frozen-model fingerprints, which must be identical:
  - M1: **`r tar_read(m1_fingerprint_check)$status`**
  - M2: **`r tar_read(m2_fingerprint_check)$status`**
  - M3: **`r if (has_m3) safe_read("m3_fingerprint_check")$status else "not frozen"`**

## B. What genuinely new information was investigated?

The full source feasibility matrix is in `docs/milestone3_plan.md` §13. In summary:

```{r b-matrix}
readr::read_csv("docs/source_feasibility.csv", show_col_types = FALSE) |> knitr::kable()
```

## C. Did new historical-valid information improve calibrated ESPN?

```{r c-historical, results = "asis"}
h <- safe_read("m3_info_comparisons")
if (is.null(h)) {
  cat("No historical-valid new-information model was evaluated (see the experiment log for the reasons).\n")
} else {
  print(knitr::kable(mutate(h, across(where(is.double), ~ round(.x, 4)))))
}
```

## D. Which M3 prospective challengers were frozen?

```{r d-registry, results = "asis"}
if (has_m3) {
  reg <- read_registry("m3")
  cat(sprintf("Frozen at **%s** (parent commit `%s`).\n\n", reg$frozen_at_utc, substr(reg$frozen_parent_commit, 1, 7)))
  purrr::iwalk(reg$models, function(m, id) {
    rule <- c(if (!is.null(m$role_rule)) sprintf("role: %s ≥ %s → %+.2f; ≤ %s → %+.2f", m$role_rule$metric,
                                                  m$role_rule$hi, m$role_rule$adj_hi, m$role_rule$lo, m$role_rule$adj_lo),
              if (!is.null(m$return_rule)) sprintf("return: 1 missed → %+.1f; 2+ missed → %+.1f",
                                                    m$return_rule$adj_one, m$return_rule$adj_two_plus))
    cat(sprintf("- **`%s`**: %s. %s\n", id, m$hypothesis, paste(rule, collapse = "; ")))
  })
  first_week <- man |> filter(grepl("m3", lineages)) |> pull(week) |> as.integer() |> min(na.rm = TRUE)
  cat(sprintf("\nProspective record starts in **week %s** (the first archived pre-kickoff run including M3).\n",
              if (is.finite(first_week)) first_week else "—"))
} else cat("No M3 challenger is frozen yet.\n")
```

One-time historical check of the frozen rules (2024–2025, rolling; not used to
change them):

```{r d-holdout, results = "asis"}
hc <- if (has_m3) safe_read("m3_holdout_comparisons") else NULL
if (!is.null(hc)) {
  print(knitr::kable(hc |> select(model, subset, n, n_weeks, mae_diff, ci_low, ci_high, weeks_better, rmse_model, rmse_baseline) |>
    mutate(across(where(is.double), ~ round(.x, 4)))))
  print(knitr::kable(safe_read("m3_holdout_seasonal") |> mutate(across(where(is.double), ~ round(.x, 4)))))
}
```

## E. What prospective evidence exists?

```{r e-prospective, results = "asis"}
if (pros$weeks_complete == 0) {
  cat("**Not enough prospective evidence yet.** No archived week has completed. Only archived, hash-verified",
      "predictions made before each game's kickoff will ever be scored. Nothing is back-filled.\n")
} else {
  print(knitr::kable(mutate(pros$metrics, across(where(is.double), ~ round(.x, 3)))))
  print(knitr::kable(mutate(pros$comparisons, across(where(is.double), ~ round(.x, 4)))))
}
```

## F. Did the role-change hypothesis work?

## G. Did the return-from-absence hypothesis work?

```{r fg, results = "asis"}
if (pros$weeks_complete < 8) {
  cat(sprintf("Prospectively: **insufficient evidence so far** (%d completed archived weeks; the pre-registered",
              pros$weeks_complete), "evaluation needs at least 8). The historical development evidence that",
      "motivated each rule, and the one-time 2024–2025 check above, are *not* confirmatory.\n")
}
```

## H. What did component modeling teach us?

```{r h-components}
cs <- tar_read(m3_component_summary)
cs$accuracy |> tidyr::pivot_wider(id_cols = component, names_from = level, values_from = mae) |> fmt()
cs$incremental |> select(component, mae_diff, ci_low, ci_high, weeks_better) |> fmt(4)
tar_read(m3_component_decomposition) |> fmt()
```

- Calibrated ESPN's points error comes from yards (about 23% of its variance),
  TDs (about 19%) and receptions (about 10%). The remainder is covariance between them.
- Our information adds a small, significant improvement only for **targets**,
  which measure opportunity. It **hurts** touchdown prediction.
- When we disagree with ESPN on a component, the realised error moves only about a
  third of the way toward us.

## I. Where does calibrated ESPN still fail?

```{r i-residuals, fig.height = 3.6, fig.width = 8}
ho <- tar_read(m2_holdout_aligned) |> filter(model == "m2_espn_cal")
ho |>
  mutate(bucket = cut(espn_proj, PROJECTION_BUCKETS, right = FALSE)) |>
  summarise(n = n(), mean_pred = mean(pred), mean_actual = mean(actual), mae = mean(abs(pred - actual)),
            large_miss = mean(abs(pred - actual) > 10), .by = bucket) |>
  arrange(bucket) |> fmt()
```

## J. Is there evidence of an incremental edge?

```{r j-verdict, results = "asis"}
cat("**Not enough prospective evidence yet.**",
    "No historical-valid new-information source improved calibrated ESPN (section C). The narrow M3 rules are",
    "frozen and are accumulating a clean prospective record. Calibrated ESPN remains the bar to beat.\n")
```
