---
title: "Milestone 2 — Incremental signal beyond ESPN"
subtitle: "Weekly WR PPR projections: development, historical holdout, and the 2026 prospective record"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(ggplot2)
suppressMessages(tar_source("R"))

cfg <- tar_read(config)
sel <- tar_read(m2_selection)
cal <- tar_read(m2_cal_choice)
dev_tab <- tar_read(m2_dev_tables)
dev_cmp <- tar_read(m2_dev_comparisons)
dev_al <- tar_read(m2_dev_aligned)
ho_tab <- tar_read(m2_holdout_tables)
ho_cmp <- tar_read(m2_holdout_comparisons)
ho_al <- tar_read(m2_holdout_aligned)
abl <- bind_rows(tar_read(m2_ablation_nf), tar_read(m2_ablation_aug))
repr <- tar_read(m2_representation)
resid <- tar_read(m2_residual_study)
leak_m2 <- tar_read(leakage_check_m2)
fp_m1 <- tar_read(m1_fingerprint_check)
fp_m2 <- tar_read(m2_fingerprint_check)

PAL <- c("#2a78d6", "#eb6834", "#1baf7a")
theme_report <- theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = "#e6e5e1", linewidth = 0.3),
        axis.text = element_text(colour = "#52514e"), plot.title.position = "plot",
        legend.position = "top", legend.justification = "left")
fmt <- function(df, digits = 3) knitr::kable(mutate(df, across(where(is.double), ~ round(.x, digits))))
cmp_view <- function(x, models, baselines, subsets = "all") {
  x |> filter(model %in% models, baseline %in% baselines, subset %in% subsets) |>
    select(model, baseline, subset, n, n_weeks, mae_diff, ci_low, ci_high, weeks_better, rmse_model, rmse_baseline)
}
pooled <- function(tab, models, subset = "all") {
  tab |> filter(.data$subset == !!subset, season == "pooled", model %in% models) |>
    select(model, n, mae, rmse, bias, rank_cor, pairwise_acc_top60, top24_precision, large_miss_rate) |>
    arrange(mae)
}
```

## Summary

- **Benchmarks:** raw ESPN, then calibrated ESPN (``r cal$model`` → `m2_espn_cal`).
  The headline comparison is **ESPN + our information (`m2_espn_aug`) vs calibrated ESPN**.
- **Protocol** (pre-registered, experiment log E3):
  - Development folds 2020–2023 with weekly expanding refits from 2018. All decisions were made there.
  - Frozen on 2026-10-05 (`models/registry/m2.yml`).
  - Historical holdout 2024–2025, run once.
  - 2026 is prospective.
- **Integrity:**
  - M2 leakage check: `r leak_m2$leaking_features` leaking features (`r leak_m2$n_cutoffs` random cutoffs, all history tables corrupted).
  - Frozen M1 fingerprint: **`r fp_m1$status`**.
  - Frozen M2 fingerprint: **`r fp_m2$status`**.

::: {.callout-important}
## Bottom line
No repeatable edge over **calibrated** ESPN has been demonstrated.

- The M2 augmentation model was marginally better on development folds, but the
  confidence interval included 0.
- It was slightly worse on the 2024–2025 holdout, and significantly worse among
  fantasy-relevant WRs.
- The only robust gain over raw ESPN is simple calibration, which carries no player-level information.
- The ESPN-free model improved modestly over M1 but remains well behind ESPN.
- The 2026 prospective record has started (Week 5 is archived) but has no completed weeks yet.
:::

## A. Did the independent (ESPN-free) model improve?

Pooled metrics on identical player-weeks (all ESPN-projected WRs).

Development folds 2020–2023:

```{r a-tables}
pooled(dev_tab, c("nf_ols", "nf_enet", "nf_xgb", "m1spec_no_espn", "naive_roll8", "espn", cal$model)) |> fmt()
```

Holdout 2024–2025:

```{r a-holdout}
pooled(ho_tab, c("m2_no_espn", "m1_no_espn", "naive_roll8", "espn", "m2_espn_cal")) |> fmt()
cmp_view(ho_cmp, "m2_no_espn", c("m1_no_espn", "m2_espn_cal"), c("all", "relevant")) |> fmt(4)
```

The M2 ESPN-free model (`m2_no_espn`, OLS on the M2 feature set) has slightly
lower MAE and clearly lower RMSE than M1's. The MAE difference is not
significant in either the development folds or the holdout. It still trails
calibrated ESPN by about 0.15–0.19 PPR points of MAE.

## B. Did we discover information beyond ESPN?

```{r b-holdout}
pooled(ho_tab, c("espn", "m2_espn_cal", "m1_espn_recal", "m1_espn_plus", "m2_espn_aug")) |> fmt()
```

**Headline: M2 ESPN + features vs calibrated ESPN.** Paired, week-clustered
bootstrap, stratified by season. Negative values favour the model.

```{r b-headline}
bind_rows(
  mutate(cmp_view(dev_cmp, sel$aug, cal$model, c("all", "relevant", "startable")), stage = "development 2020–23"),
  mutate(cmp_view(ho_cmp, "m2_espn_aug", "m2_espn_cal", c("all", "relevant", "startable")), stage = "holdout 2024–25")
) |> relocate(stage) |> fmt(4)
```

Context comparisons on the holdout:

```{r b-context}
bind_rows(
  cmp_view(ho_cmp, "m2_espn_cal", "espn"),
  cmp_view(ho_cmp, "m2_espn_aug", c("espn", "m1_espn_plus")),
  cmp_view(ho_cmp, "m1_espn_plus", "m2_espn_cal")
) |> fmt(4)
```

Calibration alone beats raw ESPN by about 0.05 MAE on the holdout. The M1
ESPN-plus model is statistically tied with calibrated ESPN. The M2 augmentation
model orders players very slightly better, but it adds positive bias and loses on
MAE.

## C. What new features actually mattered?

Descriptive ablations on static development folds. Δ = MAE without the family
minus MAE with all families, so positive means the family helped. These were
**not** used to select features.

```{r c-ablation, fig.height = 3.8, fig.width = 8}
ab <- abl |> mutate(model = recode(model, nf_ols = "ESPN-free (nf_ols)",
                                   aug_resid_enet = "ESPN augmentation (aug_resid_enet)"))
ggplot(ab, aes(delta_mae, reorder(family, delta_mae))) +
  geom_vline(xintercept = 0, colour = "#b5b4af", linewidth = 0.4) +
  geom_col(fill = PAL[1], width = 0.6) +
  facet_wrap(~model) +
  labs(title = "Change in MAE when a feature family is removed",
       x = "Δ MAE (positive = family helped)", y = NULL) +
  theme_report
```

```{r c-repr}
repr |> fmt(4)
```

For the ESPN-free model, opportunity, role-change and history features carry the
signal. On top of calibrated ESPN, no family moves MAE by more than about 0.005.
Replacing M1's trailing 3/8-game windows with exponentially weighted means plus
trend terms is the largest single representation gain (about −0.05 MAE for an
ESPN-free OLS).

## D. Did the improvement generalise historically?

```{r d-seasonal, fig.height = 3.4, fig.width = 8}
season_boot <- function(al, model, base, seasons) {
  purrr::map(seasons, function(s) {
    b <- pooled_bootstrap(filter(al, season == s), model, base, 1000, cfg$evaluation$seed)
    mutate(b, season = s)
  }) |> purrr::list_rbind()
}
sb <- bind_rows(
  mutate(season_boot(dev_al, sel$aug, cal$model, 2020:2023), stage = "development"),
  mutate(season_boot(ho_al, "m2_espn_aug", "m2_espn_cal", 2024:2025), stage = "holdout")
)
ggplot(sb, aes(factor(season), mae_diff, colour = stage)) +
  geom_hline(yintercept = 0, colour = "#b5b4af", linewidth = 0.4) +
  geom_linerange(aes(ymin = ci_low, ymax = ci_high), linewidth = 0.8) +
  geom_point(size = 2.5) +
  scale_colour_manual(values = PAL[1:2], name = NULL) +
  labs(title = "ESPN augmentation vs calibrated ESPN, by season",
       subtitle = "ΔMAE with 95% week-bootstrap interval; below 0 favours the augmentation model",
       x = NULL, y = "Δ MAE (PPR points)") +
  theme_report
sb |> select(stage, season, n, mae_diff, ci_low, ci_high, weeks_better) |> fmt(4)
```

The augmentation gain was small and positive in every development season, then
reversed in both holdout seasons. That pattern fits a weak signal partly fitted to
the development period, not a durable edge.

## E. What happened prospectively in 2026?

```{r e-prospective}
man <- readr::read_csv(PREDICTION_MANIFEST, col_types = readr::cols(.default = "c"))
ver <- verify_prediction_archive()
man |>
  left_join(select(ver, run_id, verified), by = "run_id") |>
  select(run_id, week, lineages, n_rows, predicted_at_utc, first_kickoff_utc, snapshot_captured_at_utc, git_commit, verified) |>
  mutate(git_commit = substr(git_commit, 1, 7)) |>
  knitr::kable()
off <- official_predictions(ver)
live_sched <- latest_live_file("schedules", 2026)
live_stats <- latest_live_file("player_stats", 2026)
tg26 <- clean_team_games(live_sched)
st26 <- clean_player_stats(live_stats, read_scoring_rules(cfg$scoring_system))
pros <- score_prospective(off, st26, filter(tg26, week >= 5))
```

**Completed prospective weeks: `r pros$weeks_complete`.**
`r if (pros$weeks_complete == 0) "No archived week has been played yet, so there is no prospective result. Week 5 is archived; its games begin Thursday 2026-10-08. The primary hypothesis is evaluated once, at the end of the 2026 regular season, and needs at least 10 completed archived weeks." else ""`

Prospective results only ever use archived, hash-verified predictions made before
each game's kickoff (docs/prospective_protocol.md). No backtest numbers appear in
this section.

## F. Where does the model still fail?

What calibrated ESPN misses (development folds, exploratory). Mean residual
(actual − calibrated ESPN) by group, with 95% intervals from week-level means:

```{r f-resid, fig.height = 4.2, fig.width = 8}
rs <- resid |>
  filter(factor %in% c("target_share_trend_q5", "xfp_trend_q5", "played_team_prev_game", "projection_bucket")) |>
  mutate(factor = recode(factor, target_share_trend_q5 = "Target-share trend (quintile)",
                         xfp_trend_q5 = "xFP trend (quintile)",
                         played_team_prev_game = "Played team's previous game",
                         projection_bucket = "ESPN projection range"),
         group = factor(group, levels = c("[0,5)", "[5,10)", "[10,15)", "[15,Inf)", "FALSE", "TRUE", 1:5)))
ggplot(rs, aes(mean_resid, group)) +
  geom_vline(xintercept = 0, colour = "#b5b4af", linewidth = 0.4) +
  geom_linerange(aes(xmin = ci_low, xmax = ci_high), colour = PAL[1], linewidth = 0.8) +
  geom_point(colour = "#0d366b", size = 2) +
  facet_wrap(~factor, scales = "free_y") +
  labs(title = "Where calibrated ESPN errs (development seasons)",
       subtitle = "Below 0 = over-projected", x = "Mean residual (PPR points)", y = NULL) +
  theme_report
```

- **ESPN over-reacts to recent role change.** WRs whose target share or xFP has
  recently **risen** are over-projected, and those whose share has fallen are
  under-projected.
- **Players returning from a missed game are over-projected** by about 0.6 points.
- A residual model with these features did not generalise to 2024–2025. The
  patterns may be season-specific, or too weak relative to noise.

Calibration and large misses on the holdout, by ESPN projection range:

```{r f-buckets}
ho_al |>
  filter(model %in% c("espn", "m2_espn_cal", "m2_espn_aug")) |>
  mutate(bucket = cut(espn_proj, PROJECTION_BUCKETS, right = FALSE)) |>
  summarise(n = n(), mean_pred = mean(pred), mean_actual = mean(actual), mae = mean(abs(pred - actual)),
            large_miss = mean(abs(pred - actual) > 10), .by = c(bucket, model)) |>
  arrange(bucket, model) |> fmt()
```

The largest remaining errors are boom weeks (long TDs, blow-up games) that no
pregame information anticipates. About 8% of player-weeks miss by more than 10
points under every model.

## G. Is there evidence of a repeatable edge?

**No repeatable edge: calibrated ESPN remains at least as good as our best model.**

- Pre-registered historical criterion: **not met**.
  - The development CI included 0.
  - The holdout ΔMAE was positive in both seasons.
- Prospective evidence: **none yet** (0 completed weeks).
- Positive results:
  - a measurable and replicated benefit from calibrating ESPN;
  - modest improvement of the ESPN-free model;
  - a clean, hash-committed prospective record running from Week 5;
  - exploratory evidence that ESPN over-reacts to recent role changes, a hypothesis
    to test prospectively, not a finding to act on.
